### 2022-11-07 NSF2022/Description/backgroundv2.tex

\begin{figure}
  \begin{center}
    \includegraphics[width=0.45\textwidth]{figures/SpecificAim1/Domain.jpg}
% \includegraphics[width=0.55\textwidth]{figures/SpecificAim1/Domain.jpg}    
 \includegraphics[width=0.45\textwidth]{figures/SpecificAim1/3Particles.jpg}
  \end{center}
  \caption{\label{fig:flow_map} \footnotesize {\em Left:} In the HAP
  formulation, a system of exterior problems defines the hydrodynamic
  and hydrophobic interactions. {\em Right:} Particles translate and
  rotate to lower the free energy $F[u]$ subject to \eqref{eq:SL}, $f(u)
  = u^2$, and reach a minimal configuration in panel C. The colormap is
  blue for $u = 0$ and red for $u = 1$. The arrows are the velocity
  field $\mathbf{u}$.}
\end{figure}

@@

The first key idea is to define the grain interactions using a boundary
value problem
\begin{alignat}{3}
  -\mu\Delta \uu + \nabla p &= \mathbf{0}, \:
  \nabla \cdot \uu = 0, \quad &&\xx \in \Omega_t, \label {eqn:stokes} \\
  -\rho^2 \Delta \eta + f'(\eta) &= 0, &&\xx \in \Omega_t,
  \label{eqn:phase}
\end{alignat}
with boundary conditions
\begin{alignat}{3}
  \frac{d\xx}{dt} &= \uu, &&\xx \in \bd\Omega_t \label{eqn:noslip} \\
  \uu &= \vv_i + \omega_i \times (\xx - \aa_i), \quad &&\xx \in
  \Gamma_i,\, i=1,\ldots,N_b, \label{eqn:rbm} \\
  \eta &= h_i(\xx,t),  &&\xx \in \Gamma_i,\, i=1,\ldots,N_b,
  \label{eqn:material}
\end{alignat}
and
\begin{align}
\int_{\Gamma_i} \mu(\nabla \uu + \nabla \uu^T + pI)\nu \,\dif s = 
\int_{\Gamma_i}
\underbrace{\left(-\rho^{-1} f'(\eta)
  + \rho\left(\tfrac{1}{2}|\nabla \eta|^2 - \nabla \eta \nabla
  \eta^T\right)\right)}_{T_i}
  \nnu \, \dif s.
  \label{eqn:stressbalance}
\end{align}
This is a semi-linear system for evolution of the grains.
Equation~\eqref{eqn:stokes} is the Stokes equations that describes the
velocity $\uu$ and pressure $p$ of the viscous aqueous phase.
Equation~\eqref{eqn:phase} describes the phase transitions of the scalar
order parameter $\eta$ with a decay length $\rho$. Mathematically, this
order parameter is responsible for the attraction and adhesion between
grains that want to align boundaries of like phase. Physically, water
molecules are polar and $\eta$ describes the orientational order of
water e.g.~$\eta = 1$ when water molecules share alignment and $\eta =
0$ when the water orientations are disordered. The well-potential $f$
accounts for possibly distinct equilibrium phases, as occurs with 3-bond
and 4-bond hydrogen bonding.

The next three equations are the moving domain boundary conditions.
Equations~\eqref{eqn:noslip} and~\eqref{eqn:rbm} specify a non-slip
boundary condition for a rigid body with translation velocity $\vv_i$
and angular velocity $\omega_i$ about the point $\aa_i \in \Omega_i$. In
equation~\eqref{eqn:material}, the function $h_i$ is a material label
specifying the water phase at the water-grain interface. This material
label is transported by the fluid and therefore satisfies $h_i(\xx,t) =
h^0_i(\xx(\XX,t),t)$ where $h_i^0$ and $\XX$ are reference labels,
coordinates given by the initial conditions, and $\xx(\XX,t)$ is the
flow map for the rigid bodies. Finally,
equation~\eqref{eqn:stressbalance} closes the system by requiring that
the hydrodynamic stress is balanced by the water-order stress.

Intriguingly, the
formulation~\eqref{eqn:stokes}--\eqref{eqn:stressbalance} captures
exactly the right behavior expected from amphiphilic self-assembly.
Three bodies with amphiphilic label spontaneously translate and change
orientations to sequester the disordered fluid phase. In our simulations
(see section ()), we include a finite length repulsion to avoid particle
collisions. Moreover, a simple change of boundary conditions from
amphiphilic, to biased hydrophobic, to polar results in distinct
equilibrium morphologies (Figure~\ref{fig:self-assembly}).

\begin{figure}[h]
  \begin{center}
    \begin{tabular}{m{0.1cm}m{1.8in}m{1.8in}m{1.8in}}            
      (a)
    &\includegraphics[width=1.8in]{figures/SpecificAim1/N100B1.jpg}
    &\includegraphics[width=1.8in]{figures/SpecificAim1/N100B2.jpg}
     &\includegraphics[width=1.8in]{figures/SpecificAim1/N100B3.jpg}    \\
     (b)
    &\includegraphics[width=1.8in]{figures/SpecificAim1/N100C1.jpg}
    &\includegraphics[width=1.8in]{figures/SpecificAim1/N100C2.jpg}
      &\includegraphics[width=1.8in]{figures/SpecificAim1/N100C3.jpg}    \\
    (c)
      &\includegraphics[width=1.8in]{figures/SpecificAim1/N100A1.jpg}
      &\includegraphics[width=1.8in]{figures/SpecificAim1/N100A2.jpg}
      &\includegraphics[width=1.8in]{figures/SpecificAim1/N100A3.jpg}    \\
      &\multicolumn{1}{c}{$t = 0$}
      &\multicolumn{1}{c}{intermediate time}
      &\multicolumn{1}{c}{long time}
  \end{tabular}
  \end{center}
  \vspace{-20pt}
  \caption{\footnotesize \label{fig:self-assembly} Rows (a)-(c) have the
  same initial configuration but different boundary conditions $g$. In
  row (a), self-assembly with an amphiphilic boundary condition results
  in a bilayer pattern shielding the hydrophobic core (yellow, $u > 0$)
  from the aqueous phase (blue, $u = 0$). Row (b) shows particles with a
  hydrophobic intensity that is greater on one side of the particle
  (magenta) than the other (yellow). In row (c), green is for $u < 0$,
  yellow is for $u > 0$, and blue is for $u = 0$. Particles orient like
  sides to form a checkerboard pattern.}
\end{figure}

The system~\eqref{eqn:stokes}--\eqref{eqn:stressbalance} has an elegant
mathematical structure. In terms of locally in time well-posedness,
given disjoint grains with sufficiently smooth boundaries and data, the
system always has a solution and these solutions have a locally
Lipschitz dependence on the particle centers and orientations.
Solutions of~\eqref{eqn:stokes}--\eqref{eqn:stressbalance} obey the
energy law
\begin{equation}
  \frac{d}{dt}F
  = - \int_{\Omega_t} \tfrac{1}{2}\mu|\nabla \uu + \nabla \uu^T|^2 \,d\xx,
  \quad
  F = \int_{\Omega_t}
  \rho |\nabla \eta|^2 + \rho^{-1} f(\eta) \,d\xx.
\end{equation}
Mathematically, this energy law states that the free energy $F$ of the
water order phase transitions, which depends on the grain
configurations, is lost due to hydrodynamic dissipation. It is evident
that minimizers of the functional $F$ are unique and satisfy the
Euler-Lagrange equation (). There is a subtlety in the energy law,
though. Namely, the time derivative of the free energy $F$ also involves
the moving boundary, accounted for by the Rayleigh Transport Theorem,
and the fact that the boundary data for $\eta$ are transported by the
rigid body motion. In fact, the water order surface stress $T_i$ is the
variational derivative that shows up in this calculation.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Boundary Integral formulation}
Boundary integral equations (BIEs) are a reformulation of elliptic PDEs.
For example, solutions of the Stokes equations~\eqref{eqn:stokes} can be
written as the convolution 
\begin{align*}
  \uu(\xx) = \int_{\partial\Omega_t} G(\xx - \yy) \ssigma(\yy) \, ds_\yy,
\end{align*}
where $G(\xx - \yy)$ is a carefully chosen derivative of the PDE's
fundamental solution and $\ssigma: \partial \Omega_t \rightarrow \RR^2$
is an unknown density function. The BIE reformulation has several
immediate advantages including a dimension reduction since the unknown
density function is defined only on $\partial \Omega_t$, and the far
field boundary condition of the PDE is automatically satisfied. The
density function $\ssigma$ is chosen to satisfy the boundary condition,
and this results in a BIE 
\begin{align*}
  \gg(\xx) = \lambda \ssigma(\xx) + 
    \int_{\partial\Omega_t} G(\xx - \yy) \ssigma(\yy) \, ds_\yy,
    \quad \xx \in \partial\Omega_t,
\end{align*}
where $\gg$ is the Dirichlet boundary condition, and $\lambda$ is a
constant that depends on the choice of $G$. Note that if the
well-potential is $f(u) = \tfrac{1}{2}u^2$, then
equation~\eqref{eqn:phase} becomes the so-called linear screened Laplace
equation and can be written as a BIE with an unknown scalar-valued
density function.

In addition to the dimension reduction and far field conditions being
automatically solved, additional advantages of a BIE formulation are
that they are high-order, or even spectrally, accurate; their
discretization results in a well-conditioned linear system that can be
solved iteratively with a mesh-independent number of iterations of a
Krylov method such as the generalized minimal residual method (GMRES);
and complex geometries are easier to address than mesh based methods
such as the finite element method Figure~\ref{fig:fem_vs_bie}. However,
BIEs do have several challenges: nearly touching grains requires a
quadrature rule for nearly-singular integrals; their discretization
results in dense matrices; and they are not immediately suitable to
non-linear PDEs. The BIE community, including PI Quaife, has developed a
variety of tools to address these numerical challenges, and we discuss
our planned approach in Section~\ref{sec:specificaim2}.

%The boundary integral equation (BIE) method is the ideal numerical
%method for the present study in terms of computational cost and
%accuracy. Boundary integral equations require linear PDE and so we take
%$f(\eta) = \tfrac{1}{2}\eta^2$, yielding in () the so called screened
%Laplace equation $-\rho^2 \Delta \eta + \eta$.  This is a Helmholtz
%equation with purely imaginary wave number, enabling us to represent the
%solution in terms of layer potentials:
%\begin{equation}
%  \eta(\xx) = \int_{\partial \Omega} \frac{\partial G}{\partial \nu}
%  \left(\frac{\xx-\yy}{\rho}\right) \sigma(\yy) \,ds_{\yy}
%\end{equation}
%where $\frac{\partial G}{\partial \nu}$ is the normal derivative of the
%fundamental solution of the screened Laplace equation
%(a Bessel function in two space dimensions and
%an exponential times the standard fundamental solution of the Laplace equation
%in three space dimensions).
%This representation allows us to transfer the boundary value problem ()-()
%to an equation only involving in terms of domain boundary.  
%
%The numerical advantages of doing so are as follows.  If $h = N^{-1}$ is
%the characeristic grid spacing of the grain boundaries, then a finite
%element or finite difference discretization of ()-() requires on the
%order of $N_b N^n$ grid points for dense suspensions.  In contrast BIE
%require onely With finite elements, mesh regeneration would be needed at
%each time step since the domain is moving.  In the boundary integral
%approach, collocation points can be selected for the initial grain
%configuration and then used throughout the time-course.  As a downside,
%integral equations require specialized quadrature routines to resolve
%self interactions and nearly touching grain boundaries.  Finally, the
%far field boundary conditions which are important for imposed flow
%problems automatically hold, whereas with finite element and finite
%difference methods artificial boundary conditions must be imposed.
\begin{wrapfigure}[10]{r}{2in}
  \todo[inline]{A figure comparing FEM versus BIE meshes}
  \caption{\label{fig:fem_vs_bie} \footnotesize A finite element method
  requires meshing the entire domain (typically into triangles), while a
  BIE only requires meshing the boundary of the domain.}
\end{wrapfigure}

>>>>

\begin{figure}
  \begin{center}
%    \includegraphics[width=0.45\textwidth]{figures/SpecificAim1/Domain.jpg}
%    \includegraphics[width=0.45\textwidth]{figures/SpecificAim1/3Particles.jpg}
    \includegraphics[keepaspectratio,height=2.7cm]{figures/SpecificAim1/Domain.jpg}
    \includegraphics[keepaspectratio,height=2.7cm]{figures/SpecificAim1/3Particles.jpg}
  \end{center}
  \caption{\label{fig:flow_map} \footnotesize {\em Left:} In the HAP
  formulation, a system of exterior problems defines the hydrodynamic
  and hydrophobic interactions. {\em Right:} Particles translate and
  rotate to lower the free energy $F[u]$ subject to \eqref{eq:SL}, $f(u)
  = u^2$, and reach a minimal configuration in panel C. The colormap is
  blue for $u = 0$ and red for $u = 1$. The arrows are the velocity
  field $\mathbf{u}$.}
\end{figure}

@@

The first key idea is to define the grain interactions using a boundary
value problem
\begin{alignat}{3}
  -\mu\Delta \uu + \nabla p &= \mathbf{0}, \:
  \nabla \cdot \uu = 0, \quad &&\xx \in \Omega_t, \label {eqn:stokes} \\
  -\rho^2 \Delta \eta + f'(\eta) &= 0, &&\xx \in \Omega_t,
  \label{eqn:phase}
\end{alignat}
with boundary conditions
\begin{alignat}{3}
  \frac{d\xx}{dt} &= \uu, &&\xx \in \bd\Omega_t \label{eqn:noslip} \\
  \uu &= \vv_i + \omega_i \times (\xx - \aa_i), \quad &&\xx \in
  \Gamma_i,\, i=1,\ldots,N_b, \label{eqn:rbm} \\
  \eta &= h_i(\xx,t),  &&\xx \in \Gamma_i,\, i=1,\ldots,N_b,
  \label{eqn:material}
\end{alignat}
and
\begin{align}
\int_{\Gamma_i} \mu(\nabla \uu + \nabla \uu^T + pI)\nnu \,\dif s = 
\int_{\Gamma_i}
\underbrace{\left(-\rho^{-1} f'(\eta)
  + \rho\left(\tfrac{1}{2}|\nabla \eta|^2 - \nabla \eta \nabla
  \eta^T\right)\right)}_{T_i}
  \nnu \, \dif s,
  \label{eqn:stressbalance}
\end{align}
where $\nnu$ is the unit outward normal vector. This is a semi-linear
system for evolution of the grains.  Equation~\eqref{eqn:stokes} is the
Stokes equations that describes the velocity $\uu$ and pressure $p$ of
the viscous aqueous phase.  Equation~\eqref{eqn:phase} describes the
phase transitions of the scalar order parameter $\eta$ with a decay
length $\rho$. Mathematically, this order parameter is responsible for
the attraction and adhesion between grains that want to align boundaries
of like phase. Physically, water molecules are polar and $\eta$
describes the orientational order of water e.g.~$\eta = 1$ when water
molecules share alignment and $\eta = 0$ when the water orientations are
disordered. The well-potential $f$ accounts for possibly distinct
equilibrium phases, as occurs with 3-bond and 4-bond hydrogen bonding.

The next three equations are the moving domain boundary conditions.
Equations~\eqref{eqn:noslip} and~\eqref{eqn:rbm} specify a no-slip
boundary condition for a rigid body with translation velocity $\vv_i$
and angular velocity $\omega_i$ about the point $\aa_i \in \Omega_i$. In
equation~\eqref{eqn:material}, the function $h_i$ is a material label
specifying the water phase at the water-grain interface. This material
label is transported by the fluid and therefore satisfies $h_i(\xx,t) =
h^0_i(\xx(\XX,t),t)$ where $h_i^0$ and $\XX$ are reference labels,
coordinates given by the initial conditions, and $\xx(\XX,t)$ is the
flow map for the rigid bodies. Finally,
equation~\eqref{eqn:stressbalance} closes the system by requiring that
the hydrodynamic stress is balanced by the water-order stress.

Intriguingly, the
formulation~\eqref{eqn:stokes}--\eqref{eqn:stressbalance} captures
exactly the right behavior expected from amphiphilic self-assembly.
Three bodies with amphiphilic label spontaneously translate and change
orientations to sequester the disordered fluid phase. In our simulations
(see section ()), we include a finite length repulsion to avoid particle
collisions. Moreover, a simple change of boundary conditions from
amphiphilic, to biased hydrophobic, to polar results in distinct
equilibrium morphologies (Figure~\ref{fig:self-assembly2}).

\begin{figure}[h]
  \begin{center}
    \begin{tabular}{m{0.1cm}m{1.8in}m{1.8in}m{1.8in}}            
      (a)
    &\includegraphics[width=1.8in]{figures/SpecificAim1/N100B1.jpg}
    &\includegraphics[width=1.8in]{figures/SpecificAim1/N100B2.jpg}
     &\includegraphics[width=1.8in]{figures/SpecificAim1/N100B3.jpg}    \\
     (b)
    &\includegraphics[width=1.8in]{figures/SpecificAim1/N100C1.jpg}
    &\includegraphics[width=1.8in]{figures/SpecificAim1/N100C2.jpg}
      &\includegraphics[width=1.8in]{figures/SpecificAim1/N100C3.jpg}    \\
    (c)
      &\includegraphics[width=1.8in]{figures/SpecificAim1/N100A1.jpg}
      &\includegraphics[width=1.8in]{figures/SpecificAim1/N100A2.jpg}
      &\includegraphics[width=1.8in]{figures/SpecificAim1/N100A3.jpg}    \\
      &\multicolumn{1}{c}{$t = 0$}
      &\multicolumn{1}{c}{intermediate time}
      &\multicolumn{1}{c}{long time}
  \end{tabular}
  \end{center}
  \vspace{-20pt}
  \caption{\footnotesize \label{fig:self-assembly2} Rows (a)-(c) have
  the same initial configuration but different boundary conditions $g$.
  In row (a), self-assembly with an amphiphilic boundary condition
  results in a bilayer pattern shielding the hydrophobic core (yellow,
  $u > 0$) from the aqueous phase (blue, $u = 0$). Row (b) shows
  particles with a hydrophobic intensity that is greater on one side of
  the particle (magenta) than the other (yellow). In row (c), green is
  for $u < 0$, yellow is for $u > 0$, and blue is for $u = 0$. Particles
  orient like sides to form a checkerboard pattern.}
\end{figure}

The system~\eqref{eqn:stokes}--\eqref{eqn:stressbalance} has an elegant
mathematical structure. In terms of locally in time well-posedness,
given disjoint grains with sufficiently smooth boundaries and data, the
system always has a solution and these solutions have a locally
Lipschitz dependence on the particle centers and orientations.
Solutions of~\eqref{eqn:stokes}--\eqref{eqn:stressbalance} obey the
energy law
\begin{align}
  F = \int_{\Omega_t}
  \left(\rho |\nabla \eta|^2 + \rho^{-1} f(\eta)\right) \,d\xx,
  \quad
  \frac{d}{dt}F
  = - \int_{\Omega_t} \left(\tfrac{1}{2}\mu|\nabla \uu + \nabla
  \uu^T|^2\right) \,d\xx.
\end{align}
Mathematically, this energy law states that the free energy $F$ of the
water order phase transitions, which depends on the grain
configurations, is lost due to hydrodynamic dissipation. It is evident
that minimizers of the functional $F$ are unique and satisfy the
Euler-Lagrange equation~\eqref{eqn:phase}. There is a subtlety in the
energy law, though. Namely, the time derivative of the free energy $F$
also involves the moving boundary, accounted for by the Rayleigh
Transport Theorem, and the fact that the boundary data for $\eta$ are
transported by the rigid body motion. In fact, the water order surface
stress $T_i$ is the variational derivative that shows up in this
calculation.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Boundary integral formulation}

\begin{wrapfigure}[10]{r}{2in}
  \todo[inline]{A figure comparing FEM versus BIE meshes}
  \caption{\label{fig:fem_vs_bie} \footnotesize A finite element method
  requires meshing the entire domain (typically into triangles), while a
  BIE only requires meshing the boundary of the domain.}
\end{wrapfigure}
Boundary integral equations (BIEs) are a reformulation of elliptic PDEs
including the Stokes equation~\eqref{eqn:stokes} and the phase
equation~\eqref{eqn:phase} in the special case of $f(u) =
\tfrac{1}{2}u^2$. Using the Stokes equation as an example, its solution
can be written as a convolution 
\begin{align*}
  \uu(\xx) = \int_{\partial\Omega_t} G(\xx - \yy) \ssigma(\yy) \, ds_\yy,
\end{align*}
where $G(\xx - \yy)$ is a carefully chosen derivative of the PDE's
fundamental solution and $\ssigma: \partial \Omega_t \rightarrow \RR^2$
is an unknown density function. The BIE reformulation has several
immediate advantages including a dimension reduction since $\ssigma$ is
defined only on $\partial \Omega_t$, and the far field boundary
condition of the PDE is automatically satisfied. The density
function $\ssigma$ is chosen to satisfy the boundary condition, and this
results in a BIE 
\begin{align*}
  \gg(\xx) = \lambda \ssigma(\xx) + 
    \int_{\partial\Omega_t} G(\xx - \yy) \ssigma(\yy) \, ds_\yy,
    \quad \xx \in \partial\Omega_t,
\end{align*}
where $\gg$ is the Dirichlet boundary condition, and $\lambda$ is a
constant that depends on the choice of $G$. For numerical stability,
choices of $G$ that result in second-kind Fredholm integral equations,
meaning $\lambda \neq 0$, are preferred.

A consequence of the dimension reduction in a BIE formulation is that
complex geometries are much easier to resolve than a mesh-based method
since only $\partial \Omega_t$ needs to be discretized
(Figure~\ref{fig:fem_vs_bie}). Additional advantages of a BIE
formulation includes a discretization that is high-order, or even
spectrally, accurate; results in a well-conditioned linear system that
can be solved iteratively with a mesh-independent number of iterations
of a Krylov method such as the generalized minimal residual method
(GMRES). However, BIEs do have several challenges: nearly touching
grains requires a quadrature rule for nearly-singular integrals; their
discretization results in dense matrices; and they cannot immediately be
applied to non-linear PDEs. The BIE community, including PI Quaife, has
developed a variety of tools to address these numerical challenges, and
we discuss our planned approach in Section~\ref{sec:specificaim2}.

%The boundary integral equation (BIE) method is the ideal numerical
%method for the present study in terms of computational cost and
%accuracy. Boundary integral equations require linear PDE and so we take
%$f(\eta) = \tfrac{1}{2}\eta^2$, yielding in () the so called screened
%Laplace equation $-\rho^2 \Delta \eta + \eta$. This is a Helmholtz
%equation with purely imaginary wave number, enabling us to represent the
%solution in terms of layer potentials:
%\begin{equation}
%  \eta(\xx) = \int_{\partial \Omega} \frac{\partial G}{\partial \nu}
%  \left(\frac{\xx-\yy}{\rho}\right) \sigma(\yy) \,ds_{\yy}
%\end{equation}
%where $\frac{\partial G}{\partial \nu}$ is the normal derivative of the
%fundamental solution of the screened Laplace equation
%(a Bessel function in two space dimensions and
%an exponential times the standard fundamental solution of the Laplace equation
%in three space dimensions).
%This representation allows us to transfer the boundary value problem ()-()
%to an equation only involving in terms of domain boundary.  
%
%The numerical advantages of doing so are as follows.  If $h = N^{-1}$ is
%the characeristic grid spacing of the grain boundaries, then a finite
%element or finite difference discretization of ()-() requires on the
%order of $N_b N^n$ grid points for dense suspensions.  In contrast BIE
%require onely With finite elements, mesh regeneration would be needed at
%each time step since the domain is moving.  In the boundary integral
%approach, collocation points can be selected for the initial grain
%configuration and then used throughout the time-course.  As a downside,
%integral equations require specialized quadrature routines to resolve
%self interactions and nearly touching grain boundaries.  Finally, the
%far field boundary conditions which are important for imposed flow
%problems automatically hold, whereas with finite element and finite
%difference methods artificial boundary conditions must be imposed.
